---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CopernicusDataspace <img src="man/figures/logo.svg" align="right" height="139" alt="logo" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The [Copernicus Data Space Ecosystem](https://dataspace.copernicus.eu/),
is an open ecosystem that provides free instant access to a wide range
of data and services from the Copernicus Sentinel missions and
more on our planet’s land, oceans and atmosphere. This package
provides entry points to several APIs allowing users to access
the data directly in R.

## Installation

Install latest developmental version from R-Universe:

```{r install2, eval=FALSE}
install.packages("CopernicusDataspace", repos = c('https://pepijn-devries.r-universe.dev', 'https://cloud.r-project.org'))
```

## Example

TODO

```{r example}
library(CopernicusDataspace)
## TODO
```

## Code of Conduct

Please note that the CopernicusDataspace project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
