---
output: github_document
---

```{r init, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CopernicusDataspace <img src="man/figures/logo.svg" align="right" height="139" alt="logo" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml)
[![version](https://www.r-pkg.org/badges/version/CopernicusDataspace)](https://CRAN.R-project.org/package=CopernicusDataspace)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/pepijn-devries/CopernicusDataspace/graph/badge.svg)](https://app.codecov.io/gh/pepijn-devries/CopernicusDataspace)
<!-- badges: end -->

The [Copernicus Data Space Ecosystem](https://dataspace.copernicus.eu/),
is an open ecosystem that provides free instant access to a wide range
of data and services from the Copernicus Sentinel missions and
more on our planet’s land, oceans and atmosphere. This package
provides entry points to several APIs allowing users to access
the data directly in R.

## Installation

At the moment, the package is still only experimental. It can therefore only be
installed from GitHub:

```{r install_temp, eval=FALSE}
devtools::install_github("CopernicusDataspace", username = "pepijn-devries")
```

<!-- TODO update installation procedure once available -->
<!--
Install CRAN release:

```{r install1, eval=FALSE}
install.packages("CopernicusDataspace")
```

Install latest developmental version from R-Universe:

```{r install2, eval=FALSE}
install.packages("CopernicusDataspace", repos = c('https://pepijn-devries.r-universe.dev', 'https://cloud.r-project.org'))
```
-->

## Introduction

The Copernicus Data Space Ecosystem offers access to its services through several
[APIs](https://documentation.dataspace.copernicus.eu/APIs.html). This package offers
access via the following APIs:

 * [OData](https://documentation.dataspace.copernicus.eu/APIs/OData.html);
 * [STAC](https://documentation.dataspace.copernicus.eu/APIs/STAC.html); and
 * [SentinelHub](https://documentation.dataspace.copernicus.eu/APIs/SentinelHub.html).

This README shows only essential methods for downloading data. For more detailed
information consult the respective vignettes:

 * `vignette("OData")`;
 * `vignette("STAC")`; and
 * `vignette("SentinelHub")`

## Authentication

When actually downloading files from the Copernicus Data Space Ecosystem,
you usually need some form of authentication. Examples below use credentials
stored as environment variables. For more information about how to effectively
authenticate download requests, please see `vignette("Authentication")`.

## Downloading Using STAC Catalogue

When you know the STAC identifier (`id`) and the asset name you wish
to download, you can simply use the example shown below. When you don't know
these details, you should first explore the catalogue. Please see `vignette("STAC")`
for more details on working with the STAC catalogue.

```{r download-stac, eval=TRUE, message=FALSE, fig.alt="Example of downloading with STAC API"}
library(CopernicusDataspace)
library(stars) ## For reading and plotting the downloaded file

## Only run this if an S3 secret is specified
if (dse_has_s3_secret()) {
  
  id        <- "Copernicus_DSM_COG_30_S69_00_W062_00_DEM"
  asset     <- "data"

  filename  <- dse_stac_download(id, asset, tempdir())
  tile_stac <- read_stars(filename)
  plot(tile_stac, col = hcl.colors(100), axes = TRUE)
  
}
```

## Downloading Using OData API


```{r download-odata, eval=TRUE, message=FALSE, fig.alt="Example of downloading with OData API"}
fn <- tempfile(fileext = ".dt1")

dse_odata_download_path(
  product     = "2f497806-0101-5eea-83fa-c8f68bc56b0c",
  node_path   = paste(
    "DEM1_SAR_DTE_90_20101213T034716_20130408T035028_ADS_000000_5033.DEM",
    "Copernicus_DSM_30_S09_00_E026_00", "DEM",
    "Copernicus_DSM_30_S09_00_E026_00_DEM.dt1", sep = "/"),
  destination = fn)

tile_odata <- read_stars(fn)
plot(tile_odata, col = terrain.colors(100), axes = TRUE)
```
TODO

## Downloading Using SentinelHub

TODO

## More of Copernicus

More R packages for exploring other Copernicus data services:

 * [CopernicusClimate](https://github.com/pepijn-devries/CopernicusClimate)
 Dedicated to climate change datasets
 * [CopernicusMarine](https://github.com/pepijn-devries/CopernicusMarine)
 Dedicated to marine datasets

## Code of Conduct

Please note that the CopernicusDataspace project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
