---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CopernicusDataspace <img src="man/figures/logo.svg" align="right" height="139" alt="logo" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pepijn-devries/CopernicusDataspace/actions/workflows/R-CMD-check.yaml)
[![version](https://www.r-pkg.org/badges/version/CopernicusDataspace)](https://CRAN.R-project.org/package=CopernicusDataspace)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/pepijn-devries/CopernicusDataspace/graph/badge.svg)](https://app.codecov.io/gh/pepijn-devries/CopernicusDataspace)
<!-- badges: end -->

The [Copernicus Data Space Ecosystem](https://dataspace.copernicus.eu/),
is an open ecosystem that provides free instant access to a wide range
of data and services from the Copernicus Sentinel missions and
more on our planet’s land, oceans and atmosphere. This package
provides entry points to several APIs allowing users to access
the data directly in R.

## Installation

Install latest developmental version from R-Universe:

```{r install2, eval=FALSE}
install.packages("CopernicusDataspace", repos = c('https://pepijn-devries.r-universe.dev', 'https://cloud.r-project.org'))
```

## Downloading Using STAC Catalogue

TODO

```{r download-stac, eval=TRUE, message=FALSE, fig.alt="Example of downloading with STAC API"}
library(CopernicusDataspace)
library(stars) ## For reading and plotting the downloaded file

## Only run this if an S3 secret is specified
if (dse_has_s3_secret()) {
  
  id    <- "Copernicus_DSM_COG_30_S69_00_W062_00_DEM"
  asset <- "data"

  filename <- dse_stac_download(id, asset, tempdir())
  tile <- read_stars(filename)
  plot(tile, col = hcl.colors(100), axes = TRUE)
  
}
```

## More of Copernicus

More R packages for exploring other Copernicus data services:

 * [CopernicusClimate](https://github.com/pepijn-devries/CopernicusClimate)
 Dedicated to climate change datasets
 * [CopernicusMarine](https://github.com/pepijn-devries/CopernicusMarine)
 Dedicated to marine datasets

## Code of Conduct

Please note that the CopernicusDataspace project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
