---
title: "Using the STAC Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the STAC Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In the Copernicus Data Space Ecosystem (CDSE),
[STAC stands for SpatioTemporal Asset Catalog](https://stacspec.org/en).
It is a standardized, open-source metadata specification used to structure, search,
and discover Earth Observation (EO) data.

Instead of requiring users to download massive, raw satellite images, the CDSE STAC
[RESTful API](https://en.wikipedia.org/wiki/REST) allows developers to query precise
metadata (e.g., location, time, cloud cover, and specific bands) to locate exact data
assets.

Key Functions of STAC:

  * *Data Access*: It provides a direct path (such as S3 storage links) to cloud-hosted
    imagery, enabling tools to process a subset of data without full downloads.
  * *System Interoperability*: It replaces satellite-specific extensions with a
    unified data model, allowing the same code or software to seamlessly handle diverse
    datasets like Sentinel-1 and Sentinel-2.

The package also offers features for the complementary primary catalogue via
OData. For more details on that read `vignette("OData")`.

In general to understand which STAC client the server is offering, you can
call `dse_stac_client()`.

## Data Exploration

A good starting point of exploring data with STAC is being aware which collections
of data are available in the first place. You can list them as follows:

```{r stac-collections}
library(CopernicusDataspace)
dse_stac_collections()
```

The returned `data.frame` contains descriptive information about each of the collections.
It can help you focus your search. Once you have identified a collection, you can
check which filter/search are available for further narrowing your exploration tour:

```{r queryables}
dse_stac_queryables("sentinel-1-grd") |> summary()
```

The example above shows 11 properties that can be used to focus the search.
You can start an actual search by creating a STAC search request with:
`dse_stac_search_request()`.  It creates a special
class of `httr2` request object. In essence, it is a request to the API
server, which you can modify with tidyverse operators. This sounds more
complicated than it is.

Once you have created the request, you can add tidyverse operators (like
`filter()`, `arrange()` and `slice_head()`), to modify this request.
You can join those modifications with the pipe operator (`|>` or `%>%`).
You can also query products that intersect with specific spatial features (`sf`)
using `st_intersects()`.

```{r stac-search}
library(dplyr)
library(sf)

bbox <-
  sf::st_bbox(
    c(xmin = 5.261, ymin = 52.680, xmax = 5.319, ymax = 52.715),
    crs = 4326)

dse_stac_search_request("sentinel-1-grd") |>
  filter(`sat:orbit_state` == "ascending") |>
  arrange("id") |>
  st_intersects(bbox) |>
  collect()
```

## Downloading Data

When downloading data, you could retrieve the Uniform Resource Identifier (URI) with
`dse_stac_get_uri()`. To get an URI, you need at least the asset identifier and
the specific asset. Both can be obtained with a search as shown above. The example
below shows you how:

```{r stac-uri}
dse_stac_get_uri(
  asset_id = "S2A_MSIL1C_20260109T132741_N0511_R024_T39XVL_20260109T142148",
  asset = "B01",
  collection = "sentinel-2-l1c"
)
```

This approach also needs the collection from which the asset is made available.
If you don't provide it, it will be guessed with `dse_stac_guess_collection()`.
This function is not 100% reliable, so it's best practice to provide the collection
manually. Instead of working with the URI yourself it is easier to call
`dse_stac_download()`. It will automatically takes care of required authentication
for downloading the file (if properly provided). Check `vignette("Authentication")`
for more information about the authentication process.

```{r stac-download, eval=FALSE}
dse_stac_download(
  asset_id = "S2A_MSIL1C_20260109T132741_N0511_R024_T39XVL_20260109T142148",
  asset = "B01",
  collection = "sentinel-2-l1c",
  destination = tempdir()
)
```
